---
title: "R / RStudio and the tidyverse"
author: Guillaume Coqueret
output: 
    html_notebook
---

# Introduction

## Course structure (1/2)

1. In-class sessions: mix between slides & tutorials  
2. Exercises: **practice** is the most important  
3. Project, in two steps:  
  - A short presentation of the project, due **31/10/2019**
  - The full report & code, due **15/12/2019** 
  - Please do not ask for adjournments

Questions are crucial: don't be shy, there is no such thing as a bad question.   

## Course structure (2/2)

1. Introduction to R/RStudio & the tidyverse  
2. Baseline R  
3. Plots  
4. Shiny 1 - User interface   
5. Shiny 2 - Extensions  
6. Shiny 3 - Deployment   
7. Introduction to decision trees   
8. Options!!!! => to be defined together

## Install packages

There are two steps: first you need to download the package (only once).
The packages are downloaded from servers all around the world. So we start by choosing one.
```{r CRAN_mirror}
chooseCRANmirror(ind = 60)    # Set downloading location
```

```{r packages, message = FALSE, warning = FALSE, results = "hide"}
install.packages("tidyverse") # Install (download) package: only once
install.packages("gapminder") # Install (download) package: only once
```

**NOTE**: it's easy to install packages directly in RStudio ("Tools" tab).

Then, if you want to work with it, you need to load/activate it. 
```{r load, message = FALSE, warning = FALSE, results = "hide"}
library(tidyverse)            # Activate: each time you launch RStudio
library(gapminder)            # Activate: each time you launch RStudio
```


## Comments and output

- One hashtag # precedes a comment in R.   
- Two hashtags ## precede output from a code sequence in the slides.

```{r comments}
1+1 # Test!
```

In this notebook, code will appear in grey areas (rectangles).


# Tidy data via *tidyr* 

## Instances vs variables

The **diamonds** database is included in the tidyverse. The **head**() function shows the first lines of a dataset.

```{r diamonds}
head(diamonds, 4) # The number gives the amount of rows shown
head(gapminder,3)             # Have a look!
```

One **instance** = one observation = one **row**.  
One **variable** = one **unique** characteristic = one **column**.


## Variable types 

- **number** (numerical): integer (int) or double (dbl)
- **character**: text (chr)
- **factor**: categorical, ordered (ord) or not (fct)
- **boolean**: TRUE or FALSE / **T** or **F** (bool)
- **date**: day precision (date) or second precision (time): usually starts 1970-01-01

**NOTE**: we are only concerned with rectangular / structured datasets.


## Tidy data: counter-example
The table below shows the evolution of population of countries. 
```{r rows.print = 4, cols.print = 4}
spread(gapminder[c(1:4,13:16,25:28), c(1,3,5)], key = country, value = pop)
```
   
Not tidy! The columns are not **VARIABLES**!  
(This is typically the excel format.)


## Tidy tools

The tidyverse has two functions that switch from matrix/excel format to tidy data and back:

- **gather**(): from matrix/excel to tidy data
- **spread**(): from tidy data to matrix/excel

Below, one example of data (called "not_tidy_pop"") in excel format.
<div class="columns-2">
```{r tidy_data, echo = FALSE, message = FALSE, warning = FALSE}
Year <- c(1970, 1990, 2010)
France <- c(52,59,65)
Germany <- c(61,80,82)
UK <- c(56,57,63)
not_tidy_pop <- data_frame(Year, France, Germany, UK)
not_tidy_pop
```

**BE VERY CAREFUL**: type case matters in R!     
Continent ≠ continent.   
When referring to a variable (column names), a mistake will lead to an error.


## Tidy tool: gather! From wide to long

Gather joins/concatenates columns which belong to the same variable.
```{r tidy_gather, message = FALSE, warning = FALSE}
tidy_pop <- gather(not_tidy_pop, key = Country, value = Population, -Year)
```


```{r}
tidy_pop[1:7,]  # First 7 lines (only) shown
```

The syntax is the following:

gather(**data**,    
**key** = names_of_the_new_variable,     
**value** = name_of_the_column_with_values,      
-**variable_excluded_from_the_variable**    
)

**key** = Country because the columns are all countries.

**value** = Population because the data pertains to population values.

We use -**Year** because the Year variable is excluded from the gathering.


## Tidy tools: spread! From long to wide.
The reverse operation.
```{r tidy_spread}
spread(tidy_pop, key = Country, value = Population)
```
spread(**data**,   
**key** = variable_to_be_put_in_columns,   
**value** = which_column_to_split  
)



# Data manipulation via *dplyr*

## Filter() rows - Part I
Often, analyses are performed on subsets of data. 
```{r filter()}
filter(gapminder, lifeExp > 81.5) # Countries where people live long lives on average
```


## Filter() rows - Part II 
Filters can be combined. 
```{r filter_2}
filter(gapminder, country == "Japan", year > 2000) 
```


Only two observations for Japan post-2000.  
**NOTE**: as in **all** languages, there are **TWO EQUAL SIGNS** for the comparison.  
One "=" is like the arrow (<-) and is used to **assign** values.

## Select() columns 
Sometimes, you might want to get rid of some variables to ease readability.
```{r select(), table_red}
select(gapminder[1:5,], country, year, pop)
```

## Sort via arrange() 
This is when you want to order your data. Here, from smallest pop to largest.
```{r arrange}
head(arrange(gapminder, pop)) # Alternative: arrange(gapminder, desc(lifeExp)); desc() is for descending
```

## Create new columns via mutate() 
With population and gdpPercap you can infer total GDP!
```{r mutate}
head(mutate(gapminder, gdp = pop * gdpPercap))
```

```{r}
head(gapminder)
```


# Piping: %>%

## Definition: sequences of operations 
**Very often**, one simple analysis will require several steps. They can be combined via the **%>%** operator.
```{r pipe1, results = FALSE}
select(filter(diamonds, carat > 4), carat, price, clarity) # BEURK!
```
```{r pipe2}
diamonds %>%
  select(price, clarity, carat) %>%
    filter(carat > 4) 
     # So simple!
```

## Example
```{r}
diamonds %>% 
    filter(carat > 2, cut == "Ideal") %>%      # First we filter
    mutate(car_price_ratio = carat/price) %>%  # Then, we create a new column
    arrange(desc(car_price_ratio)) %>%         # We order the data
    select(-x, -y, -z) %>%                     # We take out some superfluous columns
    head(4)                                    # Finally, we ask for the top 4 instances
```



# Pivot tables

## Definition

"*A pivot table is a table of statistics that summarizes the data of a more extensive table.*"

--- Wikipedia

There are two dimensions in a pivot table:  
- which **variable**(s) we want to analyze;  
- which **statistic** we want to compute.

In R, these two steps are separated via two functions: **group_by**() and **summarise**()

## Example I 

```{r pt1}
diamonds %>%
    group_by(clarity, cut) %>%              # Define the variables
    summarise(avg_price = mean(price),      # Define the statistics
              avg_carat = mean(carat)) %>%  # Second stats
    head()
```

## Example II

```{r pt2}
gapminder %>%
    group_by(continent, year) %>%
    summarise(avg_lifeExp = mean(lifeExp)) %>%
    head()
```


## Resources / Links

- an online series of exercises with solutions!:   
https://gcoqueret.shinyapps.io/Exercises/
- my course on statistics in R (and the many links therein!):  
https://github.com/shokru/rstats    
- Shiny use cases:   
https://www.rstudio.com/products/shiny/shiny-user-showcase/
- Other examples:   
https://www.showmeshiny.com
- My simple example of dynamic dashboard:    
https://gcoqueret.shinyapps.io/olympics/
- One other app (marketing):    
https://mdancho84.shinyapps.io/shiny-app/


